% interactnlmsample.tex
% v1.05 - August 2017

\documentclass[]{interact}
\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
\usepackage[caption=false]{subfig}% Support for small, `sub' figures and tables
\usepackage[overload]{empheq}
%\usepackage[nolists,tablesfirst]{endfloat}% To `separate' figures and tables from text if required
%\usepackage[doublespacing]{setspace}% To produce a `double spaced' document if required
%\setlength\parindent{24pt}% To increase paragraph indentation when line spacing is doubled

\usepackage[numbers,sort&compress]{natbib}% Citation support using natbib.sty
\bibpunct[, ]{[}{]}{,}{n}{,}{,}% Citation support using natbib.sty
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% Bibliography support using natbib.sty
\makeatletter% @ becomes a letter
\def\NAT@def@citea{\def\@citea{\NAT@separator}}% Suppress spaces between citations using natbib.sty
\makeatother% @ becomes a symbol again

\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newtheorem{notation}{Notation}


\usepackage{setspace}
%\usepackage[latin1]{inputenc}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{float}
\usepackage{siunitx}
%\usepackage{units}
\usepackage{listings}
\usepackage{pdfpages}
\usepackage{keyval}
\usepackage{eso-pic}
\usepackage{atbegshi}
\usepackage{pdflscape}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{url}
\usepackage{array}
\usepackage{fancyhdr}
\usepackage{enumitem}
\usepackage{lipsum}

\usepackage[np]{numprint}
\npdecimalsign{\ensuremath{.}}

\usepackage{caption}

\captionsetup[figure]{labelfont={bf},name={Fig.},labelsep=space}
\captionsetup[table]{labelfont={bf},name={Table},labelsep=space}

\newcommand {\fncyblank }{\fancyhf{}}

\newcommand{\cedpage}{\newpage{\thispagestyle{empty}\cleardoublepage}}
\usepackage{algorithm,algpseudocode}
\usepackage{tikz,tikz-3dplot}
\usetikzlibrary{arrows.meta,angles,decorations.pathmorphing,patterns}

\usepackage{bm}

\usepackage{etoolbox} % for conditional blocks

\DeclareMathOperator{\mf}{F_{\text{MF}}}

\DeclareMathOperator{\diag}{diag}

% version 1 (next 4 lines)
%\DeclareMathOperator{\VarOp}{Var}
%\DeclareMathOperator{\CovOp}{Cov}
%\newcommand\Var[1]{\VarOp(#1)}
%\newcommand\Cov[2]{\CovOp(#1,#2)}

% version 2 (next 2 lines)
\newcommand\Var[1]{\sigma_{#1}^2}
\newcommand\Cov[2]{\sigma_{#1\,#2}}


%\usepackage{hyperref}

%%% Local definitions %%%%%%
\input{mymacros}
%%% End Local definitions %%

\begin{document}

%\articletype{ARTICLE TEMPLATE}% Specify the article type or omit as appropriate

\title{Disturbance-aware minimum-time planning strategies for motorsport vehicles with probabilistic safety certificates}

\author{Martino Gulisano\textsuperscript{a}, Matteo Masoni\textsuperscript{a}, \name{Marco Gabiccini\textsuperscript{a,*} and Massimo Guiggiani\textsuperscript{a}
\thanks{* Contact: Marco Gabiccini. Email: marco.gabiccini@unipi.it}
}\affil{
\textsuperscript{a} Dipartimento di Ingegneria Civile e Industriale,	Universit\`{a} di Pisa, Pisa, Italy.}
}

\maketitle

\begin{abstract}
This paper presents a disturbance-aware framework that embeds robustness into minimum-lap-time trajectory optimization for motorsport. Two formulations are introduced. (i) Open-loop, horizon-based covariance propagation uses worst-case uncertainty growth over a finite window to tighten tire-friction and track-limit constraints. (ii) Closed-loop, covariance-aware planning incorporates a time-varying LQR feedback law in the optimizer, providing a feedback-consistent estimate of disturbance attenuation and enabling sharper yet reliable constraint tightening. Both methods yield reference trajectories for human or artificial drivers: in autonomous applications the modelled controller can replicate the on-board implementation, while for human driving accuracy increases with the extent to which the driver can be approximated by the assumed time-varying LQR policy.  Computational tests on a representative Barcelona-Catalunya sector show that both schemes meet the prescribed safety probability, yet the closed-loop variant incurs smaller lap-time penalties than the more conservative open-loop solution, while the nominal (non-robust) trajectory remains infeasible under the same uncertainties.
By accounting for uncertainty growth and feedback action during planning, the proposed framework delivers trajectories that are both performance-optimal and probabilistically safe, advancing minimum-time optimization toward real-world deployment in high-performance motorsport and autonomous racing.
\end{abstract}

\begin{keywords}
Minimum lap-time trajectory planning; stochastic vehicle dynamics; probabilistic safety certificates.
\end{keywords}



%\tableofcontents
\input{intro}
\input{stochastic_vehicle_dynamics}
\input{open_loop_planning}
%for back-off calculations
\input{closed_loop_planning}
\input{performance_comparison}
\input{conclusions}

\section*{Acknowledgement(s)}

The authors would like to thank Dr. Lorenzo Bartali for his significant contributions to the initial version of the optimization framework developed and utilized in this paper.


\section*{Disclosure statement}

No potential conflict of interest was reported by the author(s).

\section*{Funding}

This work is supported by project PRIN 2022 PNRR ``Global Stability of road vehicle motion - STAVE'' CUP I53D23005670001.



%\section*{Notes on contributor(s)}
%
%An unnumbered section, e.g.\ \verb"\section*{Notes on contributors}", may be included \emph{in the non-anonymous version} if required. A photograph may be added if requested.
%
%
%\section*{Nomenclature/Notation}
%
%An unnumbered section, e.g.\ \verb"\section*{Nomenclature}" (or \verb"\section*{Notation}"), may be included if required, before any Notes or References.
%
%
%\section*{Notes}
%
%An unnumbered `Notes' section may be included before the References (if using the \verb"endnotes" package, use the command \verb"\theendnotes" where the notes are to appear, instead of creating a \verb"\section*").


\bibliographystyle{tfnlm}
\bibliography{Disturbance-Aware-Planning}


%\bigskip
%
%
%\section{Appendices}
%
%\input{Appendices}

%\section{Troubleshooting}




%\section{Obtaining the template and class file}
%
%\subsection{Via the Taylor \& Francis website}
%
%...
%
%
%\subsection{Via e-mail}

\end{document}
